Fast and scalable Gaussian process modeling with applications to astronomical time series \vspace{-3\baselineskip}

### 2017-01-12 paper/ms.tex
adding rotation figures

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.9\textwidth]{figures/astero-11615890-comparisons.pdf}
\caption{A comparison between the empirical PSD and the posterior inference of
the PSD as a mixture of stochastically driven simple harmonic oscillators.
(top) The periodogram of the \kepler\ light curve for KIC~11615890 computed
    on the full four year baseline of the mission.
    The light gray curve shows the raw periodogram and the black curve has
    been smoothed with a Gaussian filter with \dfmtodo{SOME WIDTH}.
(middle) The same periodogram computed using about a month of data.
(bottom) The inferred power spectrum under the mixture of SHOs model described
    in the text.
    The black line shows the median of posterior PSD and the gray contours
    show the 68\% credible region.
    \figurelabel{astero}}
\end{center}
\end{figure}

@@

For many stars it is possible

>>>>

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.9\textwidth]{figures/astero-11615890-comparisons.pdf}
\caption{A comparison between the empirical PSD and the posterior inference of
the PSD as a mixture of stochastically driven simple harmonic oscillators.
(top) The periodogram of the \kepler\ light curve for KIC~11615890 computed
    on the full four year baseline of the mission.
    The light gray curve shows the raw periodogram and the black curve has
    been smoothed with a Gaussian filter with \dfmtodo{SOME WIDTH}.
(middle) The same periodogram computed using about a month of data.
(bottom) The power spectrum inferred using the mixture of SHOs model described
    in the text and only one month of \kepler\ data.
    The black line shows the median of posterior PSD and the gray contours
    show the 68\% credible region.
    \figurelabel{astero}}
\end{center}
\end{figure}

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/astero-11615890-numax_deltanu_corner.pdf}
\caption{The probabilistic constraints on $\nu_\mathrm{max}$ and $\Delta \nu$
    from the inference shown in \Figure{astero} compared to the published
    value (error bar) based on the full \kepler\ dataset \dfmtodo{CITE}.
    \figurelabel{astero-corner}}
\end{center}
\end{figure}

@@

Another source of variability that can be measured from time series
measurements of stars is rotation.
The inhomogeneous surface of the star (spots, plage, \etc) imprints itself as
quasiperiodic variations in photometric or spectroscopic observations
(\dfmtodo{CITE}).
It has been demonstrated that for light curves with nearly uniform sampling,
the empirical autocorrelation function provides a reliable estimate of the
rotation period of a star (\dfmtodo{CITE}) and that a GP model with a
quasiperiodic covariance function can be used to make probabilistic
measurements even with sparsely sampled data (R.~Angus, \etal\ in prep.).
The covariance function used for this type of analysis has the form
\begin{eqnarray}\eqlabel{sine2}
k(\tau) = A\,\exp\left(-\frac{\tau^2}{2\,\ell^2} -
    \Gamma\,\sin^2\left(\frac{\pi\,\tau}{P} \right) \right)
\end{eqnarray}
where $P$ is the period of the oscillation.
The key difference between this function and other quasiperiodic kernels is
that it is everywhere positive.
We can construct a simple \genrp\ covariance function with similar properties
as follows
\begin{eqnarray}\eqlabel{rot-kernel}
k(\tau) = \frac{a}{2+b}\,e^{-c\,\tau}\,\left[
    \cos\left(\frac{2\,\pi\,\tau}{P}\right) + (1 + b)
\right]
\end{eqnarray}
for $a>0$, $b>0$, and $c>0$.
The covariance function in \eq{rot-kernel} cannot exactly reproduce \eq{sine2}
but, since \eq{sine2} is only an effective model, \eq{rot-kernel} can be used
as a drop-in replacement for a substantial gain in computational efficiency.

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=\textwidth]{figures/rotation.pdf}
\caption{Inferred constraints on a quasiperiodic GP model using the covariance
    function in \eq{rot-kernel} and one quarter of \kepler\ data.
(left) The \kepler\ data (black points) and the maximum likelihood model
    prediction (blue curve).
    The solid blue line shows the predictive mean and the blue contours show
    the predictive standard deviation.
(center) Inferred constraints on the model PSD.
    The dashed line shows the maximum likelihood PSD, the blue solid line
    shows the median of posterior PSD, and the blue contours show the 68\%
    credible region.
(right) Inferred constraints on the model covariance function from
    \eq{rot-kernel}.
    The dashed line shows the maximum likelihood model, the blue solid line
    shows the median of posterior, and the blue contours show the 68\%
    credible region.
    \figurelabel{rotation}}
\end{center}
\end{figure}

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.5\textwidth]{figures/rotation-period.pdf}
\caption{The posterior constraint on the rotation period of KIC~1430163 using
    the dataset and model from \Figure{rotation}.
    The period is the parameter $P$ in \eq{rot-kernel} and this figure shows
    the posterior distribution marginalized over all other nuisance parameters
    in \eq{rot-kernel}.
    This is consistent with the published rotation period made using the
    autocorrelation function and the full \kepler\ data (\dfmtodo{CITE}).
    \figurelabel{rotation-period}}
\end{center}
\end{figure}
